---
title: "Data Wrangling"
description: |
  In this document, we retreive and wrangle the NMMAPS City Data.
author:
  - name: Vincent Bagilet 
    url: https://www.vincentbagilet.github.io/
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
date: "May 11, 2021"
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

```{r setup_wrangling, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/data_wrangling/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               eval = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

```{r package_wrangling, echo=TRUE, message=FALSE, warning=FALSE, code_folding='Show code to load packages'}
library(rmarkdown)
library(knitr) 
library(here) 
library(tidyverse) 
library(lubridate) 
```

# Raw data

To access the raw data, we use the `NMMAPSdata` R package and developed by Roger D. Peng and Leah J. Welty that contains the data of the National Morbidity, Mortality, and Air Pollution Study (NMMAPS). Unfortunately, the package is not longer available on CRAN but its archives are extremely [well-documented](https://journal.r-project.org/articles/RN-2004-011/RN-2004-011.pdf) and allow to easily rebuild the dataset. We store all the raw `.rda` data, city by city in a folder. 

# Loading the raw data

First we create a dataframe with paths to the data.

```{r paths_data_wrangling, message = FALSE, warning = FALSE}
folder_raw_data <- here::here("data", "raw_data", "city_data")

city_data <- tibble(file_path = list.files(path = folder_raw_data,
                        pattern = ".rda", 
                        full.names = F)) %>%
  mutate(city = str_remove(string = file_path, pattern = ".rda"))
```

Then, we write two functions that load and  wrangle the data, `load_rda` and `wrangle_city_data` respectively.

```{r load_wrangle_wrangling}
load_rda <- function(fileName) {
  #loads an RData file, and returns it
  load(fileName)
  get(ls()[ls() != "fileName"])
}

wrangle_city_data <- function(file_path){
  data <- load_rda(here::here("data", "raw_data", "city_data", file_path)) %>%
      select(
        date,
        agecat,
        accident,
        copd,
        cvd,
        death,
        inf,
        pneinf,
        pneu,
        resp,
        tmean,
        pm10mean,
        pm25mean,
        o3mean,
        so2mean,
        no2mean,
        comean
      ) %>%
      rename("temperature" = tmean) %>%
      # convert temperature to celcius degrees
      mutate(temperature = (temperature-32)/1.8) %>%
      mutate(
        agecat = case_when(
          agecat == 1 ~ "age_below_65",
          agecat == 2 ~ "age_65_75",
          agecat == 3 ~ "age_above_75"
        ),
        date = lubridate::ymd(date)
      )
    
    names(data) <- str_remove_all(names(data), "mean")
    
    # reshape to wide by age categories
    data <- data %>%
      pivot_wider(names_from = agecat,
                  values_from = c(accident:resp)) %>%
      mutate(
        accident_total = rowSums(select(., starts_with("accident"))),
        copd_total = rowSums(select(., starts_with("copd"))),
        cvd_total = rowSums(select(., starts_with("cvd"))),
        death_total = rowSums(select(., starts_with("death"))),
        inf_total = rowSums(select(., starts_with("inf"))),
        pneinf_total = rowSums(select(., starts_with("pneinf"))),
        pneu_total = rowSums(select(., starts_with("pneu"))),
        resp_total = rowSums(select(., starts_with("resp")))
      )
  
  return(data)
} 
```

We then apply this function to each file path.

```{r create_city_data}
city_data <- city_data %>%
  mutate(data = map(file_path, wrangle_city_data)) %>%
  select(-file_path) %>%
  unnest(cols = c(data))
```

# Adding variables

We then add useful variables to the dataset. First, we start with city-level metadata stored in the "codebook" folder.

```{r metadata_wrangling}
metadata_city <- load_rda(here::here("data", "raw_data", "codebook", "cities.rda")) %>%
  select(city, cityname, state, statename) %>%
  mutate(cityname = paste0(cityname, paste0(", ", state))) %>%
  select(-state) %>%
  rename("city_name" = cityname, "state" = statename)
  
city_data <- left_join(city_data, metadata_city, by = "city") %>%
  select(-city) %>%
  rename(city = city_name)
```

We then add calendar variables (weekday, month and year).

```{r calendar_wrangling}
city_data <- city_data %>%
  mutate(weekday = wday(date, label = TRUE, abbr = FALSE),
         month = month(date, label = TRUE, abbr = FALSE),
         year = as.factor(year(date)))
```

Finally, we fix negative values of $CO$ and convert its concentrations from $ppb$ to $\mu g.m^{-3}$

```{r fix_co_wrangling}
city_data <- city_data %>%
  mutate(
    co = ifelse(co < 0, abs(co), co),
    co = 1.145*co
  )
```

# Save data

We can now save the data after reordering the variables.

```{r save_data_wrangling, echo=TRUE}

# save data
city_data <- city_data %>%
  select(city, state, date, weekday, month, year, temperature:co,
         accident_age_below_65, accident_age_65_75, accident_age_above_75, 
         accident_total,
         copd_age_below_65, copd_age_65_75, copd_age_above_75, copd_total,
         cvd_age_below_65, cvd_age_65_75, cvd_age_above_75, cvd_total,
         death_age_below_65, death_age_65_75, death_age_above_75, death_total,
         inf_age_below_65, inf_age_65_75, inf_age_above_75, inf_total,
         pneinf_age_below_65, pneu_age_65_75, pneu_age_above_75, pneu_total,
         resp_age_below_65, resp_age_65_75, resp_age_above_75, resp_total) %>%
  saveRDS(., here::here("data", "clean_data", "nmmaps_data.rds"))
```





