---
title: "Data Cleaning for Power Simulations"
description: |
  In this document, we select NMMAPS cities with the least proportion of missing records to use them in our power simulations.
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
date: "May 11, 2021"
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

```{r setup_cleaning, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/data_cleaning/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

```{r package_cleaning, echo=TRUE, message=FALSE, warning=FALSE, code_folding='Show code to load packages'}
library(rmarkdown) 
library(knitr) 
library(here) 
library(tidyverse)
library(missRanger) 
library(mediocrethemes) #own ggplot theme

set_mediocre_all(pal = "leo")
```

# Selecting Cities with the Least Proportion of Missing Observations

First, we load the data. We then notice in the exploratory data analysis that average temperature readings were missing for all cities after 1998. We therefore select observations from 1987 to 1997:

```{r dates_cleaning, echo=TRUE, message = FALSE, warning = FALSE}
nmmaps_to_clean <- readRDS(here::here("data", "clean_data", "nmmaps_data.rds"))

nmmaps_filtered <- nmmaps_to_clean %>%
  filter(date <= "1997-12-31")
```

We select cities with less than 5% of missing temperature readings:

```{r temperature_cleaning, message = FALSE, warning = FALSE}
cities_to_keep <- nmmaps_filtered %>%
  group_by(city) %>%
  summarise(proportion_missing = sum(is.na(temperature))/n()*100) %>%
  filter(proportion_missing<=5) %>%
  pull(city)

nmmaps_filtered <- nmmaps_filtered %>%
  filter(city %in% cities_to_keep)
```

We will only use the carbon monoxide in our power simulations as it is the pollutant with the smallest proportion of missing readings. We keep cities with less than 5% of missing CO readings:

```{r co_cleaning, message = FALSE, warning = FALSE}
cities_to_keep <- nmmaps_to_clean %>%
  group_by(city) %>%
  summarise(proportion_missing = sum(is.na(co))/n()*100) %>%
  filter(proportion_missing<=5) %>%
  pull(city)

nmmaps_to_clean <- nmmaps_to_clean %>%
  filter(city %in% cities_to_keep)
```

We have `r length(unique(nmmaps_to_clean$city))` cities remaining. We plot the proportion of missing values for CO by city:

```{r missing_prop_cleaning, message = FALSE, warning = FALSE, fig.width=8, fig.height=20, layout="l-body-outset", code_folding=TRUE}
nmmaps_to_clean %>%
  group_by(city) %>%
  summarise(proportion_missing = sum(is.na(co))/n()*100) %>%
  arrange(proportion_missing) %>%
  mutate(city = factor(city, levels = city)) %>%
  ggplot(., aes(x = proportion_missing, y = city)) +
  geom_vline(xintercept = sum(is.na(nmmaps_to_clean$co))/nrow(nmmaps_to_clean)*100) +
  geom_point(size = 3) +
  xlab("Proportion of Missing Values (%)") + 
  ylab(NULL) 
```

We also plot the time series of missing values for CO by city:

```{r missing_city_cleaning, message = FALSE, warning = FALSE, fig.width=8, fig.height=20, layout="l-body-outset", code_folding=TRUE}
nmmaps_to_clean %>%
  mutate(missing_co = ifelse(is.na(co), "Missing", "Not Missing")) %>%
  ggplot(., aes(x = date, y = city, fill = fct_rev(missing_co))) +
  geom_tile() +
  scale_x_date(breaks = scales::pretty_breaks(n = 6)) +
  labs(
    x = "Date",
    y = NULL,
    fill = NULL
  ) + 
  theme(panel.grid.major.y = element_blank(),
        legend.justification = "left",
        axis.text.y = element_text(margin = margin(r = -0.5, unit = "cm")))
```

The overall distribution of CO in $\mu g/m^{3}$ is as follows:

```{r distrib_co_cleaning, warning=FALSE, code_folding=TRUE}
nmmaps_to_clean %>%
  summarise(Mean = mean(co, na.rm = TRUE),
            "Standard Deviation" = sd(co, na.rm = TRUE),
            Minimum = min(co, na.rm = TRUE),
            Maximum = max(co, na.rm = TRUE)) %>%
  kable(.)
````

We impute missing observations for CO concentrations and the average temperature using the `missRanger` package:

```{r imputation_cleaning, echo=FALSE, warning=FALSE, eval=FALSE}
nmmaps_imputed <- missRanger::missRanger(
  nmmaps_to_clean,
  # variables to impute
  co +
  temperature ~
  # variables used for the imputation
  .,
  pmm.k = 10,
  num.trees = 100
)

saveRDS(nmmaps_imputed, here::here("data", "clean_data", "nmmaps_imputed.rds"))
````

We scale ("standardize") the CO concentrations:

```{r scale_cleaning, echo=TRUE, warning=FALSE, eval=FALSE}
nmmaps_data_simulations <- nmmaps_imputed %>%
  mutate(co = scale(co) %>% as.vector)
````

We finally keep relevant variables and save the data:

```{r save_cleaning, echo=TRUE, message = FALSE, warning = FALSE, eval=FALSE}
nmmaps_data_simulations %>%
  select(city:temperature, co:resp_total) %>%
  mutate_at(vars(city, state, year), ~ as.factor(.)) %>%
  saveRDS(., here::here("data", "clean_data", "nmmaps_data_simulations.rds"))
```

```{r load_final_cleaning, echo=FALSE}
nmmaps_data_simulations <- 
  readRDS(here::here("data", "clean_data", "nmmaps_data_simulations.rds"))
```

We have `r nrow(nmmaps_data_simulations)` observations, with `r nrow(nmmaps_data_simulations)/length(unique(nmmaps_data_simulations$city))` daily observations by city. 
