---
title: "Exaggeration in experimental economics"
description: "In this document, leveraging data form @camerer_evaluating_2016, I examine potential exaggeration in experimental economics using replication results as hypothetical true effect sizes. The experimental literature seems to display less exaggeration on average that non-experimental ones but this figure hides substantial heterogeneity, exaggeration being large for some studies."
output: 
  distill::distill_article:
    toc: true
    float: true
# bibliography: paper/current_version/air_pollution.bib
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup_comparison, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/lit_exp/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 300,
               fig.align = "center",
               dev.args = list(bg="transparent"))  
```  

```{r packages_comparison, message=FALSE, warning=FALSE, code_folding='Show the packages used in this document'}
library(tidyverse) 
library(knitr) 
library(mediocrethemes)
library(here)
library(retrodesign)
library(haven)
library(DT)
library(kableExtra)

set_mediocre_all(pal = "leo")
```

## Setting and data

@camerer_evaluating_2016 replicate 18 experimental economics studies published in the *American Economic Review* and the *Quarterly Journal of Economics* between 2011 and 2014. They find that original studies on average overestimate the replicated effect by a factor of 1.5 and only 61% of the replication find a significant effect in the same direction as the original study.

**In this document, I report and replicate their results and further explore exaggeration in the original studies.** This data allows to have a convincing estimate of the true effect (the replicated point estimate) and to evaluate the ability of the original design to accurately capture this estimate. 

I retrieve the data from the [project website](http://experimentaleconreplications.com/scripts.html). I just ran their Stata script `create_studydetails.do` to generate their data set. Since the standard errors of the estimates are not reported in this data set, I recompute them based on a normal distribution.

```{r load_data_camerer, code_folding=TRUE}
data_camerer_raw <- read_dta("data/lit_review_econ/studydetails.dta") 

data_camerer <- data_camerer_raw |>  
  mutate(
    se_orig = (eorigu95 - eorig)/qnorm(0.975),
    se_rep = (erepu95 - erep)/qnorm(0.975),
  ) |> 
  select(
    study, 
    p_value_orig = porig, 
    p_value_rep = prep, 
    se_orig, 
    se_rep, 
    estimate_orig = eorig, 
    estimate_rep = erep
  ) |> 
  mutate(
    ratio = abs(estimate_orig/estimate_rep),
    same_sign_signif = 
      (sign(estimate_orig) == sign(estimate_rep)
             & p_value_rep < 0.05)
  )
```

## Replicability and Ratio Origingal/Replicated Estimate

I first reproduce the results in @camerer_evaluating_2016, evaluating the proportion of studies that find results in the same direction as the original study. I also compute the median ratio of the original over the replicated estimate.

```{r replicability, code_folding=TRUE}
data_camerer |> 
  mutate(same_sign_and_signif = ifelse(same_sign_signif, "Yes", "No")) |> 
  group_by(same_sign_and_signif) %>%
  summarise(
    number = n(),
    proportion = n()/nrow(.),
    `median_ratio_orig/rep` = median(ratio),
    `3rd_quartile_ratio_orig/rep` = quantile(ratio, 0.75)
  ) |>
  rename_with(\(x) str_to_title(str_replace_all(x, "_", " "))) |>
  kable(digits = 2) 
```

The ratios of the original over the replicated estimate do not seem to be too large for studies that replicate but are more concerning for those that do not replicate:

```{r distrib_ratio, code_folding=TRUE}
data_camerer |> 
  # filter(same_sign_and_signif) |> 
  arrange(desc(ratio)) |>  
  mutate(
    same_sign = ifelse(sign(estimate_orig) == sign(estimate_rep), "✓", " "),
    signif_rep = ifelse(p_value_rep < 0.05, "✓", " "),
    signif_orig = ifelse(p_value_orig < 0.05, "✓", " ")
  ) |> 
  select(
    study_id = study, 
    `ratio_orig/rep` = ratio, 
    same_sign, 
    signif_rep,
    signif_orig
  ) |>
  rename_with(\(x) str_to_title(str_replace_all(x, "_", " "))) |>
  kable(digits = 2, align = "c") 
```


## Power Calculations

### Power and Exaggeration

I then compute the statistical power and exaggeration of the initial analysis. To do so, I assume that the true effect is equal to the effect found in the replication and compute the corresponding statistical power and exaggeration of the original study using the `retrodesign` package.

```{r camerer_retro}
retro_camerer <- data_camerer |> 
  mutate(
    retro = map2(estimate_rep, se_orig, \(x, y) retro_design_closed_form(x, y))
    #retro_design returns a list with power, type_s, type_m
  ) |> 
  unnest_wider(retro) |> 
  mutate(power = power * 100, type_s = type_s * 100)
```

For most studies, power and exaggeration are limited but some of them display limitations:

```{r median_exagg, code_folding=TRUE}
source(here("functions.R"))

retro_camerer |> 
  summary_power() |>
  rename_with(\(x) str_to_title(str_replace_all(x, "_", " "))) |>
  kable()
```

Plotting the whole distribution of exaggeration ratio and power clarifies this heterogeneity:

```{r distrib_camerer_retro, code_folding=TRUE}
retro_camerer |> 
  filter(p_value_orig < 0.05) |> 
  ggplot() +
  geom_dotplot(aes(x = type_m), color = NA) +
  labs(
    title = "Distribution of the exaggeration ratio in the original studies",
    subtitle = "If the true effect was equal to the replicated one",
    x = "Exaggeration ratio (log scale)",
    y = "Number of studies",
    caption = "Each dot represents one study"
  ) + 
  scale_x_log10() +
  scale_y_continuous(NULL, breaks = NULL) 

retro_camerer |> 
  ggplot() +
  geom_dotplot(aes(x = power), color = NA) +
  labs(
    title = "Distribution of the power in the original studies",
    subtitle = "If the true effect was equal to the replicated one",
    x = "Power (%)",
    y = "Number of studies",
    caption = "Each dot represents one study"
  ) +
  scale_y_continuous(NULL, breaks = NULL) 
```

A non-negligible portion of the studies has a low power to detect the original results and are therefore likely to produce inflated statistically significant estimates. One study produces extremely large exaggeration.

### Adequate Power

I finally compute the proportion of original studies that would have adequate power as defined by the customary and arbitrary 80% threshold, still assuming that the true effect is equal to the replication one.

```{r adequate_camerer, code_folding=TRUE}
retro_camerer |> 
  group_by(
    adequate_power = ifelse(power >= 80, "Yes", "No")
  ) %>%
  summarise(
    nb = n(),
    prop = n()/nrow(.)
  ) |> 
  kable(col.names = c(
    "Adequate power", 
    "Number",
    "Proportion"),
    digits = 2
  )
```

All these results show that even the experimental literature suffers from power and exaggeration issues, despite power being central to this literature. 

<!-- ## Randomized Control Trials -->

<!-- We want to look at replications of RCTs in Development Economics. To do so, we use [the list of replication papers put together by Sandip Sukhtankar](https://www.aeaweb.org/articles?id=10.1257/aer.p20171120). -->

<!-- We gather the list of RCTs that have been replicated in Development Economics. -->

<!-- ```{r rep_dvpt} -->
<!-- rep_dvpt_raw <- read_dta(here("Misc", "replication_data_final.dta")) -->

<!-- rep_dvpt <- rep_dvpt_raw |>  -->
<!--   filter( (Replicated == "Replicated")) -->

<!-- rep_dvpt |> count(Title) -->
<!-- ``` -->






