---
title: "Retrospective Analysis of the Epidemiology Literature Using Meta-Analyses Results"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache.path = "cache/",
  cache = FALSE,
  fig.path = "images/meta_analysis_epi/",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  out.width = "85%",
  dpi = 300,
  fig.align = "center"
)
```

In this document, we explore the relationship between the reported estimated effect sizes of air pollutants on health outcomes and the precision of these estimates in the epidemiology literature. Since our automated search of the literature does not allow us to easily standardize estimates, we rely on two recently published meta-analysis by [Orellano et al. (2020)](https://www.sciencedirect.com/science/article/pii/S0160412020318316#s0120) and [Shah et al. (2015)](https://www.bmj.com/content/350/bmj.h1295) whose data are openly available:

* [Orellano et al. (2020)](https://www.sciencedirect.com/science/article/pii/S0160412020318316#s0120) selected 196 studies on the short-term associations between air pollutants and mortality for all causes but also for specific causes. Their data are available in the Supplementary Files they provided.
* [Shah et al. (2015)](https://www.bmj.com/content/350/bmj.h1295) gathered 94 studies on the short-term associations between air pollutants and admission to hospital or mortality due to stroke. Their data are available in the "Appendix six - Individual forest plots for each pollutant" of the  [Data Supplement](https://www.bmj.com/content/350/bmj.h1295/related#datasupp).

Compared to the causal inference literature on this topic, the "standard" epidemiology literature is much more homogeneous since it has produced many studies with very similar designs (i.e., same statistical model, same specification, same health outcome, etc.).

For each study of these two datasets, we also compute their statistical power, type M and S errors using the results of the meta-analysis as guesses for the true effect sizes of an air pollutant on an health outcome.

# Loading Packages and Data

```{r loading_packages, code_folding="Show the packages used", results='hide'}
library("groundhog")
packages <- c(
  "here",
  "tidyverse", 
  "knitr",
  "retrodesign",
  "janitor",
  "mediocrethemes"
  # "vincentbagilet/mediocrethemes"
)

# groundhog.library(packages, "2022-11-28")
lapply(packages, library, character.only = TRUE)

set_mediocre_all(pal = "leo")
```

We open the two meta-analysis datasets and clean them.

```{r clean_data, code_folding=TRUE}
data_orellano_2020 <-
  read_csv(here::here(
    "data",
    "meta_analyses_epi",
    "meta_analysis_orellano_2020.csv"
  )) |> 
  janitor::clean_names() |> 
  # remove studies using the one hour maximum concentration of no2
  mutate(time_period = ifelse(is.na(time_period), "unknown", time_period)) |> 
  filter(time_period != "1 hr") |> 
  select(
    article, 
    continent, 
    age_group, 
    sex, 
    pollutant, 
    unit_measurement = unit_of_measurement,
    time_period,
    cause_death = cause_of_death,
    point_estimate = "single_pollutant_effect_estimate",
    lower_bound_ci = "single_pollutant_lower_interval_limit_95_percent_ci",
    upper_bound_ci = "single_pollutant_higher_interval_limit_95_percent_ci"
  ) |> 
  mutate(year = str_extract(article, "[[:digit:]]+") %>% as.numeric(.))

data_shah_2015 <-
  readxl::read_excel(here::here("data", "meta_analyses_epi", "meta_analysis_shah_2015.xlsx")) %>%
  mutate(air_pollutant = str_to_upper(air_pollutant))
```


# Orellano et al. (2020) Meta-Analysis

We start by exploring the studies gathered by [Orellano et al. (2020)](https://www.sciencedirect.com/science/article/pii/S0160412020318316#s0120) on the short-term associations between air pollutants and mortality for all causes and specific causes. We first select studies focusing mortality for all ages and sexes. We then compute our metric for precision, which the inverse of the standard error, and also calculate the width of the 95% confidence intervals:

```{r}
# select estimates for all ages and sexes
data_orellano_2020 <- data_orellano_2020 %>%
  filter(age_group == "All") %>%
  filter(sex == "Both")

# compute standard error and precision
data_orellano_2020 <- data_orellano_2020 %>%
  mutate(
    ci_width = upper_bound_ci - lower_bound_ci,
    standard_error = (upper_bound_ci - point_estimate) / 1.96,
    precision = 1 / standard_error
  )
```

### Relative Risks versus Precision

We plot below the relationship between estimated relative risks for all-cause mortality and precision for each air pollutant (N=`r nrow(data_orellano_2020 %>% filter(cause_death == "All-cause"))`  estimates):

```{r, fig.asp=0.6}
data_orellano_2020 %>%
  filter(cause_death == "All-cause") %>%
  ggplot(., aes(x = precision, y = point_estimate)) +
  geom_hline(yintercept = 1, colour = "black", size = 0.25, linetype = "solid") +
  geom_point(alpha = 0.8) +
  scale_x_log10() +
  facet_wrap(~ pollutant, ncol = 4) +
  labs(x = "Precision (Inverse of the Standard Error)", y = "Estimated Relative Risk", title = "Relative Risks vs. Precision for All-Cause Mortality")
```

We draw the same initial graph but for respiratory mortality (N=`r nrow(data_orellano_2020 %>% filter(cause_death == "Respiratory"))` estimates):

```{r, fig.asp=0.6}
data_orellano_2020 %>%
  filter(cause_death == "Respiratory") %>%
  ggplot(., aes(x = precision, y = point_estimate)) +
  geom_hline(yintercept = 1, colour = "black", size = 0.25, linetype = "solid") +
  geom_point(alpha = 0.8) +
  scale_x_log10() +
  facet_wrap(~ pollutant, ncol = 2) +
  labs("Precision (Inverse of the Standard Error)", y = "Estimated Relative Risk", title = "Relative Risks vs. Precision for Respiratory Mortality")
```

We draw the same previous graph but now for cardiovascular mortality (N=`r nrow(data_orellano_2020 %>% filter(cause_death == "Cardiovascular"))` estimates):

```{r, fig.asp=0.6}
data_orellano_2020 %>%
  filter(cause_death == "Cardiovascular") %>%
  ggplot(., aes(x = precision, y = point_estimate)) +
  geom_hline(yintercept = 1, colour = "black", size = 0.25, linetype = "solid") +
  geom_point(alpha = 0.8) +
  scale_x_log10() +
  facet_wrap( ~ pollutant, scales = "free", ncol = 4) +
  labs(x = "Precision (Inverse of the Standard Error)", y = "Estimated Relative Risk", title = "Relative Risks vs. Precision for Cardiovascular Mortality")
```


### Evolution of Precision over Time

We plot below the mean width of 95% confidence intervals over time for estimated relative risks of all-cause mortality:

```{r, fig.asp=0.5}
data_orellano_2020 %>%
  filter(cause_death == "All-cause") %>%
  group_by(pollutant, year) %>%
  summarise(mean_ci_width = mean(ci_width)) %>%
  ggplot(., aes(x = year, y = mean_ci_width)) +
  geom_line(linetype = "dashed") +
  geom_point() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 3)) +
  facet_wrap(~ pollutant, ncol = 4) +
  labs(x = "Year", y = "Average Width of 95%\nConfidence Intervals", title = "Evolution of the Width of Confidence Intevals Over Time")
```

### 95% CI Width as a Percentage of Estimated Relative Risk

We finally display the distribution of the 95% CI width expressed as percentage of the estimated relative risk. 

```{r}
data_orellano_2020 %>%
  filter(cause_death == "All-cause") %>%
  mutate(ci_width_percentage_effect_size = ci_width / point_estimate * 100) %>%
  group_by(pollutant) %>%
  summarise(
    Mean = mean(ci_width_percentage_effect_size),
    "Standard Deviation" = sd(ci_width_percentage_effect_size),
    Minum = min(ci_width_percentage_effect_size),
    Maximum = max(ci_width_percentage_effect_size)
  ) %>%
  mutate_at(vars(Mean:Maximum), ~ round(., 1) %>% paste(., "%")) %>%
  rename(Pollutant = pollutant) %>%
  kable(., align = "lcccc")
```

Results on PM$_{10}$ are the more precise.

### Computing Statistical Power, Type M and S errors

We compute in this section the statistical power, the probability to make a Type S error and the exaggeration factor (the average Type M error) for studies included in the meta-analysis. As a guess for the true effect size of each study, we extract the meta-analysis estimates displayed in the Table 1 of the article. We then rely on the [https://github.com/andytimm/retrodesign](https://github.com/andytimm/retrodesign) package to compute the statistical power, type M and S error. 

Two important remarks:
* We had to convert the estimates as percentage increases because the `retrodesign()` function does not work with relative risks (assuming a linear relationship, [Orellano et al. (2020)](https://www.sciencedirect.com/science/article/pii/S0160412020318316#s0120) use the following formula: Percentage Increase = (RR-1)$\times$100).
* We only consider studies that were initially deemed statistically significant at the 5% level.

We display below the median of the three metrics by air pollutant:
  
```{r}
data_retrodesign_orellano_2020 <- data_orellano_2020 %>%
  # select statistically significant estimates
  mutate(statistically_significant = ifelse(lower_bound_ci > 1, "True", "False")) %>%
  filter(statistically_significant == "True") %>%
  # convert standard error as percentage increase
  mutate(standard_error = standard_error * 100) %>%
  # add meta-analysis estimates
  mutate(
    true_effect_size = case_when(
      # meta-estimates for all-cause mortality
      pollutant == "PM10" & cause_death == "All-cause" ~ 0.41,
      pollutant == "PM2.5" & cause_death == "All-cause" ~ 0.65,
      pollutant == "O3" & cause_death == "All-cause" ~ 0.43,
      pollutant == "NO2" & cause_death == "All-cause" ~ 0.72,
      # meta-estimates for cardiovascular mortality
      pollutant == "PM10" &
        cause_death == "Cardiovascular" ~ 0.6,
      pollutant == "PM2.5" &
        cause_death == "Cardiovascular" ~ 0.92,
      # meta-estimates for respiratory mortality
      pollutant == "PM10" & cause_death == "Respiratory" ~ 0.91,
      pollutant == "PM2.5" &
        cause_death == "Respiratory" ~ 0.92,
      # meta-estimates for cerebrovascular mortality
      pollutant == "PM10" &
        cause_death == "Cerebrovascular" ~ 0.44,
      pollutant == "PM2.5" &
        cause_death == "Cerebrovascular" ~ 0.72
    )
  )

# compute power, type m and s errors
data_retrodesign_orellano_2020 <- data_retrodesign_orellano_2020 |> 
  mutate(
    retro = map2(true_effect_size, standard_error, 
                 \(x, y) retro_design_closed_form(x, y))
    #retro_design returns a list with power, type_s, type_m
  ) |> 
  unnest_wider(retro) |> 
  mutate(power = power * 100, type_s = type_s * 100)

# display median of metrics by air pollutant
data_retrodesign_orellano_2020 %>%
  group_by(pollutant, cause_death) %>%
  summarise(
    "Number of Studies" = n(),
    "Statistical Power (%)" = median(power) %>% round(., 0),
    "Type M Error" = median(type_m) %>% round(., 0),
    "Type S Error (%)" = median(type_s) %>% round(., 0)
  ) %>%
  rename("Air Pollutant" = pollutant, "Cause of Death" = cause_death) %>%
  ungroup() %>%
  arrange(`Statistical Power (%)`) %>%
  kable(., align = c("l", rep("c", 5)))
```

# Shah et al. (2015) Meta-Analysis

In this section, we analyse the studies gathered by Shah et al. (2015) on the link between short-term exposure to air pollution and stroke hospital admission or mortality. As for the previous dataset, we compute our measure of precision and the width of 95% confidence intervals:

```{r}
# compute precision and width of 95% ci
data_shah_2015 <- data_shah_2015 %>%
  mutate(
    precision = 1 / standard_error,
    upper_bound_ci = estimate + 1.96 * standard_error,
    lower_bound_ci = estimate - 1.96 * standard_error,
    ci_width = upper_bound_ci - lower_bound_ci
  )
```

### Relative Risks versus Precision

We plot below the relationship between estimated relative risks and precision for each air pollutant:

```{r, fig.asp=0.8}
data_shah_2015 %>%
  ggplot(., aes(x = precision, y = estimate)) +
  geom_hline(yintercept = 1, colour = "black", size = 0.25, linetype = "solid") +
  geom_point(alpha = 0.8) +
  scale_x_log10(n.breaks = 5) +
  facet_wrap(~ air_pollutant, ncol = 3) +
  labs(x = "Precision (Inverse of the Standard Error)", y = "Estimated Relative Risk", title = "Relative Risks vs. Precision")
```

We plot the same figure but for estimated relative risks below 2:

```{r, fig.asp=0.8}
data_shah_2015 %>%
  filter(estimate < 2) %>%
  ggplot(., aes(x = precision, y = estimate)) +
  geom_hline(yintercept = 1, colour = "black", size = 0.25, linetype = "solid") +
  geom_point(alpha = 0.8) +
  scale_x_log10(n.breaks = 5) +
  facet_wrap(~ air_pollutant, ncol = 3) +
  labs(x = "Precision (Inverse of the Standard Error)", y = "Estimated Relative Risk", title = "Relative Risks vs. Precision")
```

### Evolution of Precision over Time

We plot below the mean width of 95% confidence intervals over time for estimated relative risks:

```{r, fig.asp=0.8}
data_shah_2015 %>%
  group_by(air_pollutant, year) %>%
  summarise(mean_ci_width = mean(ci_width)) %>%
  ggplot(., aes(x = year, y = mean_ci_width)) +
  geom_line(linetype = "dashed") +
  geom_point() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 2)) +
  facet_wrap( ~ air_pollutant, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = "Average Width of 95%\nConfidence Intervals", title = "Evolution of the Width of Confidence Intevals Over Time")
```

### 95% CI Width as a Percentage of Estimated Relative Risk

We finally display the distribution of the 95% CI width expressed as percentage of the estimated relative risk. 

```{r}
data_shah_2015 %>%
  mutate(ci_width_percentage_effect_size = ci_width / estimate * 100) %>%
  group_by(air_pollutant) %>%
  summarise(
    Mean = mean(ci_width_percentage_effect_size),
    "Standard Deviation" = sd(ci_width_percentage_effect_size),
    Minum = min(ci_width_percentage_effect_size),
    Maximum = max(ci_width_percentage_effect_size)
  ) %>%
  mutate_at(vars(Mean:Maximum), ~ round(., 1) %>% paste(., "%")) %>%
  rename(Pollutant = air_pollutant) %>%
  kable(., align = "lcccc")
```


### Computing Statistical Power, Type M and S errors

As a guess for the true effect size of each study, we extract the meta-analysis estimates displayed in the Figure 1 of the article. We then rely on the [https://github.com/andytimm/retrodesign](https://github.com/andytimm/retrodesign) package to compute the statistical power, type M and S error. 

We display below the median of the three metrics by air pollutant:
  
```{r}
data_shah_2015 <- data_shah_2015 %>%
  # select statistically significant estimates
  mutate(statistically_significant = ifelse(lower_bound_ci > 1, "True", "False")) %>%
  filter(statistically_significant == "True") %>%
  # convert standard error as percentage increase
  mutate(standard_error = standard_error * 100) %>%
  # add meta-analysis estimates
  mutate(
    true_effect_size = case_when(
      air_pollutant == "PM10" ~ 0.3,
      air_pollutant == "PM2.5" ~ 1.1,
      air_pollutant == "O3" ~ 0.1,
      air_pollutant == "NO2" ~ 1.4,
      air_pollutant == "SO2" ~ 1.9,
      air_pollutant == "CO" ~ 1.5
    )
  )

# compute power, type m and s errors
data_shah_2015 <- data_shah_2015 |> 
  mutate(
    retro = map2(true_effect_size, standard_error, 
                 \(x, y) retro_design_closed_form(x, y))
    #retro_design returns a list with power, type_s, type_m
  ) |> 
  unnest_wider(retro) |> 
  mutate(power = power * 100, type_s = type_s * 100)

# display median of metrics by air pollutant
data_shah_2015 %>%
  group_by(air_pollutant) %>%
  summarise(
    "Statistical Power (%)" = mean(power) %>% round(., 0),
    "Type M Error" = mean(type_m) %>% round(., 0),
    "Type S Error (%)" = mean(type_s) %>% round(., 0)
  ) %>%
  rename("Air Pollutant" = air_pollutant) %>%
  kable(., align = "lccc")
```

We display below the distribution of the exaggeration ratio by air pollutant:

```{r, fig.asp=0.8}
# store median type m errors
median_type_m <- data_shah_2015 %>%
  group_by(air_pollutant) %>%
  summarise(median = median(type_m))
  

# make the graph
graph_shah_2015 <- data_shah_2015 %>%
  ggplot() +
  # geom_rect(aes(xmin = 1, xmax = 12, ymin = -0.5, ymax = 0.5), fill = "#dee2e6", alpha = 0.05) +
  geom_linerange(aes(x = type_m, ymin = -0.5, ymax = 0.5), colour = "#0081a7", alpha = 0.6) +
  geom_linerange(
    data = median_type_m,
    aes(x = median, ymin = -0.5, ymax = 0.5),
    size = 1.1,
    colour = "#f07167"
  ) +
  facet_wrap(~ air_pollutant) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(1, 12)) +
  labs(x = "Exaggeration Factor", y = NULL, title = NULL) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

# display the graph
graph_shah_2015

# save graph
ggsave(
  graph_shah_2015,
  filename = here::here(
    "images",
    "graph_shah_2015.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm"
  # device = cairo_pdf
)
``` 




