---
title: "Analysis of power simulations"
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
  - name: Léo Zabrocki 
    url: https://www.parisschoolofeconomics.eu/en/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
runtime: shiny
# output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

## Purpose of the document

In this document, we analyze the results of our simulations.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) 
library(readr)
library(mediocrethemes)
library(shiny)
library(here)
library(shinythemes)
library(shinyWidgets)
library(ggridges)

set_mediocre_all()

# sim_param_base <- readRDS("data/sim_param_base.RDS")
source(here("R/Shiny/functions_shiny.R"))

# summary_simulations <- readRDS(here("R/Outputs/summary_simulations.RDS"))
```

## Overall results

First of all, we built a Shiny app to analyze all our results. This app was helpful considering the large number of output variables and parameters considered. 

<iframe src="https://vincentbagilet.shinyapps.io/shiny_air_poll/" class="l-screen-inset shaded" height="1000px"></iframe>

## Graphs in the paper

In this section, we display the code used to generate the graphs in the paper. 

### Evolution of type M with the number of days

```{r graph_typeM_ndays}
baseline_n_days <- filter_baseline_var(summary_evol_small, "n_days", "type_m")  %>% 
  bind_rows(filter_baseline_var(summary_evol_large, "n_days", "type_m"))

baseline_n_days %>% 
  mutate(
    id_method = ifelse(id_method == "reduced_form", "Reduced form", id_method)
  ) %>% 
  ggplot(aes(x = n_days, y = type_m, color = as.factor(n_cities))) + #, color = .data[[id_method]] + 
  geom_point() +
  geom_line(linetype = "dashed", size = 0.1) +
  facet_wrap(~ id_method, scales = "free_x") +
  ylim(c(0, 8)) +
  labs(
    # title =  "Evolution of power with the number of days",
    # subtitle = "Comparison across identification methods",
    x = "Number of days in the study",
    y = "Type M error",
    color = "Number of cities"
  ) 

# ggsave(here("R/Graphs/evol_typeM_n_days.png"), dpi = 700, width = 9, height = 6)
```

### Evolution of power with the proportion of treated units

```{r graph_power_ptreat}
baseline_power_p_treat <- filter_baseline_var(summary_evol_small, "p_obs_treat", "power")  %>% 
  bind_rows(filter_baseline_var(summary_evol_large, "p_obs_treat", "power"))

baseline_power_p_treat %>% 
  mutate(
    id_method = ifelse(id_method == "reduced_form", "Reduced form", id_method)
  ) %>% 
  ggplot(aes(x = p_obs_treat, y = power, color = as.factor(n_cities*n_days))) + #, color = .data[[id_method]] + 
  geom_point() +
  geom_line(linetype = "dashed", size = 0.1) +
  facet_wrap(~ id_method, scales = "free_x") +
  ylim(c(0, 100)) +
  labs(
    # title =  "Evolution of power with the number of days",
    # subtitle = "Comparison across identification methods",
    x = "Proportion of treated units",
    y = "Power",
    color = "Number of observations"
  ) 

# ggsave(here("R/Graphs/evol_power_p_treat.png"), dpi = 700, width = 9, height = 4)
```

### Evolution of type M with the proportion of treated units

```{r graph_typeM_ptreat}
baseline_typeM_p_treat <- filter_baseline_var(summary_evol_small, "p_obs_treat", "type_m")  %>% 
  bind_rows(filter_baseline_var(summary_evol_large, "p_obs_treat", "type_m"))

baseline_typeM_p_treat %>% 
  mutate(
    id_method = ifelse(id_method == "reduced_form", "Reduced form", id_method)
  ) %>% 
  ggplot(aes(x = p_obs_treat, y = type_m, color = as.factor(n_cities*n_days))) + #, color = .data[[id_method]] + 
  geom_point() +
  geom_line(linetype = "dashed", size = 0.1) +
  facet_wrap(~ id_method, scales = "free_x") +
  ylim(c(0, NA)) +
  labs(
    # title =  "Evolution of power with the number of days",
    # subtitle = "Comparison across identification methods",
    x = "Proportion of treated units",
    y = "Type M error",
    color = "Number of observations"
  ) 

# ggsave(here("R/Graphs/evol_typeM_p_treat.png"), dpi = 700, width = 9, height = 4)
```

### Evolution of type M with IV strength

```{r graph_typeM_IVstrength}
baseline_iv <- filter_baseline_var(summary_evol_small, "iv_strength", "type_m")  %>% 
  bind_rows(filter_baseline_var(summary_evol_large, "iv_strength", "type_m"))

baseline_iv %>% 
  pivot_longer(cols = c(power, type_m, mean_f_stat), names_to = "stat") %>% 
  mutate(
    stat = case_when(
      stat == "mean_f_stat" ~ "F-stat",
      stat == "type_m" ~ "Type M error",
      stat == "power" ~ "Power",)
  ) %>% 
  ggplot(aes(x = iv_strength, y = value, color = as.factor(n_cities*n_days))) + #, color = .data[[id_method]] + 
  geom_point() +
  geom_line(linetype = "dashed", size = 0.1) +
  facet_wrap(~ fct_rev(stat), scales = "free_y", ncol = 1) +
  # ylim(c(0, NA)) +
  labs(
    # title =  "Evolution of power with the number of days",
    # subtitle = "Comparison across identification methods",
    x = "Strength of the IV",
    y = "",
    color = "N obs"
  ) 

# ggsave(here("R/Graphs/evol_iv_strength.png"), dpi = 700, width = 3, height = 7)
```

<!-- ```{r graph_typeM_fstat} -->
<!-- baseline_iv %>%  -->
<!--   # filter(mean_f_stat < 5000) %>% -->
<!--   filter(type_m < 10) %>%  -->
<!--   ggplot(aes(x = mean_f_stat, y = type_m)) + -->
<!--   geom_point() + -->
<!--   geom_line(linetype = "dashed", size = 0.1) + -->
<!--   ylim(c(0, NA)) -->
<!-- ``` -->


<!-- ```{r, echo = FALSE} -->
<!-- shinyAppDir( -->
<!--   here("R/Shiny/power_simulations_viz"), -->
<!--   options = list(width = "100%", height = 1300) -->
<!-- ) -->
<!-- ``` -->

<!-- ## Quick analysis -->

<!-- First, we graph of the evolution of type M, type S and power for each quasi-experiment as a function of different parameters, holding other parameters constant. This function takes as input a "summary_simulations" type of data frame. -->

<!-- ```{r} -->
<!-- graph_evol_by_exp <- function(df, var_param = "n_days_study", stat = "power") { -->

<!--   var_param_name <- str_replace_all(var_param, "_", " ") -->
<!--   stat_name <- str_replace_all(stat, "_", " ") -->

<!--   #considering baseline values -->
<!--   df_filtered <- df %>%  -->
<!--     filter(str_detect(formula, "temperature")) #to only consider the model with all covariates -->

<!--   if (var_param != "p_treat") { -->
<!--     df_filtered <- df_filtered %>%  -->
<!--       filter(p_treat == sim_param_base[["p_treat"]]) -->
<!--   }  -->
<!--   if (!(var_param %in% c("n_days_study", "average_n_obs"))) { -->
<!--     df_filtered <- df_filtered %>%  -->
<!--       filter(n_days_study == sim_param_base[["n_days_study"]]) -->
<!--   }  -->
<!--   if (var_param != "percent_effect_size") { -->
<!--     df_filtered <- df_filtered %>%  -->
<!--       filter(percent_effect_size == sim_param_base[["percent_effect_size"]]) -->
<!--   } -->

<!--   #graph itself -->
<!--   graph <- df_filtered %>%  -->
<!--     mutate( -->
<!--       quasi_exp = str_to_sentence(str_replace_all(quasi_exp, "_", " ")) -->
<!--     ) %>%  -->
<!--     ggplot(aes(x = .data[[var_param]], y = .data[[stat]])) + #, color = .data[[quasi_exp]] +  -->
<!--     geom_point() + -->
<!--     geom_line(linetype = "dashed", size = 0.1) + -->
<!--     facet_wrap(~ quasi_exp) + -->
<!--     ylim(c(0, ifelse(stat == "power", 100, NA))) + -->
<!--     labs( -->
<!--       title = paste( -->
<!--         str_to_title(stat_name), ifelse(stat == "power", "increases", "decreases"), -->
<!--         "with", var_param_name -->
<!--       ), -->
<!--       subtitle = "Comparison across quasi-experiments", -->
<!--       x = var_param_name, -->
<!--       y = str_to_title(stat_name) -->
<!--     )  -->
<!--     # theme(legend.position = "none") -->

<!--   return(graph) -->
<!-- }  -->

<!-- # graph_evol_by_exp(summary_simulations) -->
<!-- ``` -->

<!-- We then plot all the graphs. To do so, we create a tibble containing every  parameter and statistics we want to plot and map our function on this tibble. -->

<!-- ```{r} -->
<!-- stat <- c("power", "type_m", "type_s") -->
<!-- var_param <-  c("n_days_study", "average_n_obs", "percent_effect_size", "p_treat") -->

<!-- param_by_exp <- crossing(stat, var_param) -->

<!-- pmap(param_by_exp, graph_evol_by_exp, df = summary_simulations) -->
<!-- ``` -->

<!-- We can also plot the graphs for each parameter separately, for a better readability. -->

<!-- ### Length study -->

<!-- ```{r} -->
<!-- map( -->
<!--   stat,  -->
<!--   graph_evol_by_exp,  -->
<!--   df = summary_simulations,  -->
<!--   var_param = "n_days_study" -->
<!-- ) -->

<!-- map( -->
<!--   stat,  -->
<!--   graph_evol_by_exp,  -->
<!--   df = summary_simulations,  -->
<!--   var_param = "average_n_obs" -->
<!-- ) -->
<!-- ``` -->

<!-- ### Effect size -->

<!-- ```{r} -->
<!-- map( -->
<!--   stat,  -->
<!--   graph_evol_by_exp,  -->
<!--   df = summary_simulations,  -->
<!--   var_param = "percent_effect_size" -->
<!-- ) -->
<!-- ``` -->

<!-- ### Proportion of treated days -->

<!-- ```{r} -->
<!-- map( -->
<!--   stat,  -->
<!--   graph_evol_by_exp,  -->
<!--   df = summary_simulations,  -->
<!--   var_param = "p_treat" -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # check_distrib_estimate <- function(df) { -->
<!-- #  -->
<!-- #   #only consider baseline values -->
<!-- #   df_baseline <- df %>%  -->
<!-- #     filter(str_detect(formula, "resp_total")) %>%  -->
<!-- #     select(quasi_exp, n_days, n_cities, p_obs_treat, percent_effect_size, id_method) %>%  -->
<!-- #     inner_join( -->
<!-- #       df, -->
<!-- #       by = c("quasi_exp", "n_days", "n_cities", "p_obs_treat", "percent_effect_size", "id_method") -->
<!-- #     ) %>%  -->
<!-- #     filter(str_detect(formula, "death_total")) -->
<!-- #  -->
<!-- #   graph <- df_baseline %>% -->
<!-- #     # mutate(significant = (p_value < 0.05)) %>% -->
<!-- #     filter(id_method == "OLS") %>% -->
<!-- #     ggplot() + -->
<!-- #     geom_density(aes(x = estimate)) + -->
<!-- #     geom_vline(aes(xintercept = mean(true_effect))) + -->
<!-- #     # facet_wrap(~ id_method) -->
<!-- #     labs( -->
<!-- #       title = "Distribution of estimates by identification method", -->
<!-- #       subtitle = "Comparison to the true effect" -->
<!-- #     ) -->
<!-- #      -->
<!-- #   return(graph) -->
<!-- # } -->

<!-- ``` -->


