---
title: "Creation of missing observations"
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
  - name: Léo Zabrocki 
    url: https://www.parisschoolofeconomics.eu/en/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
# output:
#   html_notebook:
#     toc: no
#     theme: simplex
#     highlight: pygments
#   html_document:
#     theme: simplex
#     highlight: pygments
# ---
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

```{r include=FALSE}
library(tidyverse)
library(lubridate)
```

```{r include=FALSE}
#data wrangling copied/pasted from missingness_pattern_fr.Rmd
data_daily <- readRDS("../Outputs/data_daily.RDS") %>% 
  mutate(
    year = as.factor(year(date)),
    month = as.factor(month(date)),
    day_of_month = as.factor(day(date)),
    day_in_week = as.factor(wday(date)),
    missing = is.na(concentration)
  ) %>% 
  # filter(year != 2014) %>% 
  # filter(!(year == 2015 & pollutant == "no")) %>% 
  # filter(!(year == 2019 & str_sub(pollutant, 1, 2) == "pm")) 
# For "new" long dataset (including 2008-2012)
  filter(year != 2012) %>%
  filter(!(year %in% c(2008:2010, 2014, 2015, 2019, 2020) & pollutant == "no")) %>%
  filter(!(year %in% c(2008:2010, 2019) & str_sub(pollutant, 1, 2) == "pm")) %>% 
  select(-DIR) 
```

In this document, we build functions to create missingness patterns. We also apply them to our daily data set, to illustrate. We chose the option to create a logical variable `missing_crea` instead of deleting the observations to offer the possibility to keep the observation but to solely make the concentration missing (with a simple `mutate(concentration = ifelse(missing_crea, NA, concentration)`).

# Data missing completely at random

Here, we just create randomly missing observations.

```{r}
random_sample <- function(n, prop) {
  missing <- rep(FALSE, n)
  missing[sample(1:n, n*prop)] <- TRUE
  return(missing)
}

CAR <- data_daily %>% 
  mutate(missing_crea = random_sample(n(), 0.1)) 
```

We could also determine whether a distribution is missing or not by drawing from a Bernoulli distribution.

```{r}
CAR_bern <- data_daily %>% 
  mutate(missing_crea = rbernoulli(n(), p = 0.1)) 
```

The first method is slightly better because it returns a proportion of missing data slightly closest to the proportion required (this difference may be more important for small samples): 

```{r}
CAR %>% 
  count(missing_crea) %>% 
  mutate(prop = n/sum(n)) %>% 
  kable()

CAR_bern %>% 
  count(missing_crea) %>% 
  mutate(prop = n/sum(n))%>% 
  kable()
```

To check the performance of our methods, we create a simple function which verifies the proportion of missing values.

```{r}
prop_missing <- function (df, var = "missing_crea") {
  p <- df %>% 
    count(.data[[var]]) %>% 
    mutate(prop = n/sum(n))
  
  return(p)
} 
```

To directly drop the observations, we could have used the following code:

```{r}
CAR_drop <- data_daily %>% 
  slice_sample(prop = 1 - 0.1)
```

# Accounting for length of missingness periods

When data is missing, it is often missing for several days in a row. We want to model this.

## A unique length

Here, we consider a unique length of missingness (`length_period`). To do so, we create a function. It randomly selects dates for which data is missing and set the #`length_period` following periods to also be missing. Note that this function may create longer periods of missingness than what specified in `length_period` if the random selection selects to observations which are away from less than the value of `length_period`. The lowest the proportion of missing data, the less likely this will happen.

```{r}
unique_missing_length <- function(prop_missing, length_period) {
  draw <- random_sample(n(), prop_missing/length_period)*length_period
  # draw <- rbernoulli(n(), p = prop_missing/length_period)*length_period
  out <- draw
  
  for (i in seq_along(draw)) {
    if(draw[i] > 0) {
      for (k in 1:draw[i]) {
          out[i + k - 1] <- draw[i]
      }
    }
  }
  
  missing <- (out > 0)
  
  return(missing[1:length(draw)])
}
```

Example of implementation:

```{r}
one_length <- data_daily %>%  
  group_by(city, pollutant) %>%
  mutate(missing_crea = unique_missing_length(0.2, 4)) %>% 
  ungroup()
```

## Various lengths

In the real data set, the length of missingness periods varies. We want to account for this. We first investigate the length of missing periods (per city-pollutant and per pollutant).

```{r}
#copied/pasted from missingness_pattern_fr.Rmd
#computes the length of missingness periods
length_missing_data <- data_daily %>% 
  mutate(row_id = row_number()) %>% 
  group_by(site, pollutant) %>%
  arrange(date) %>% 
  mutate(
    missing_period_id = ifelse(missing == TRUE & lag(missing) == FALSE, row_id, NA), 
    missing_period_id = ifelse(missing == FALSE, "m", missing_period_id)
  ) %>%
  # filter(missing == TRUE) %>%
  fill(missing_period_id) %>% 
  ungroup() %>% 
  select(-row_id) %>% 
  arrange(missing_period_id) %>% 
  group_by(missing_period_id, pollutant) %>% 
  mutate(
    length_period_missing = ifelse(missing_period_id == "m", 0, n())
  ) %>% 
  ungroup()

#summarises the length of missingness periods by city-pollutant
length_per_city_poll <- length_missing_data %>% 
  filter(length_period_missing > 0) %>% 
  group_by(city, pollutant) %>% 
  count(length_period_missing) %>% 
  mutate(
    n = n/length_period_missing, 
    prop = n/sum(n)
  ) %>% 
  select(-n) %>% 
  nest() %>%
  rename(length_prop = data) %>% 
  ungroup()

#summarises the length of missingness periods by pollutant
length_per_poll <- length_missing_data %>% 
  filter(length_period_missing > 0) %>% 
  group_by(pollutant) %>% 
  count(length_period_missing) %>% 
  mutate(
    n = n/length_period_missing, 
    prop = n/sum(n)
  ) %>% 
  select(-n) %>% 
  nest() %>%
  rename(length_prop = data) %>% 
  ungroup()
```

Then, we create our function. The idea is that, for each length of missing period, it creates the demanded set of periods, independently from already existing missing observations. This can lead some periods to be overlaying each other. Again, the lowest the proportion of missing data, the less likely this will happen.

Note that, if necessary, I can modify the function to limit this.

```{r}
various_missing_length <- function(prop_missing, df_prop) {
  missing <- rep(0, n())
  df_prop_uniq <- df_prop[[1]]
  null_list <- sapply(df_prop, is_list)
  
  if(null_list[1]) {#to avoid issue when df_prop is a NULL list (due to merging)
    for (i in rev(1:nrow(df_prop_uniq))){
      missing <- missing +
        unique_missing_length(prop_missing*df_prop_uniq[[i, "prop"]],
                          df_prop_uniq[[i, "length_period_missing"]])
    }
    out <- (missing > 0)
  } else {
    out <- rep(FALSE, n())
  }
  
  return(out)
}
```

Examples: 

```{r}
missing_per_city_poll <- data_daily %>% 
  left_join(length_per_city_poll, by = c("city", "pollutant")) %>% 
  select(date, city, pollutant, length_prop) %>% 
  group_by(city, pollutant) %>% 
  mutate(missing_crea = various_missing_length(0.1, length_prop)) %>% 
  select(-length_prop) %>% 
  ungroup()

missing_per_poll <- data_daily %>% 
  left_join(length_per_poll, by = "pollutant") %>% 
  select(date, pollutant, length_prop) %>% 
  group_by(pollutant) %>% 
  mutate(missing_crea = various_missing_length(0.1, length_prop)) %>% 
  select(-length_prop) %>% 
  ungroup()
```

Note that, in `missing_per_city_poll`, the proportion of missing data is often lower than required, probably because some "long" periods of missingness are interrupted by the end of the series. (To be verified)

# One observation every 6 days

To reproduce the US sampling process, we consider only one observation every six day. To do so, we group the data by groups of 6 days and only consider the first observation of each group. To avoid considering the same day for each city-pollutant, we offset the groups by a random number.

```{r}
one_obs_per_6 <- data_daily %>% 
  group_by(city, pollutant) %>% 
  mutate(rand = sample(1:6, 1)) %>% 
  ungroup() %>% 
  mutate(six_days_period = (as.integer(date) + rand) %/% 6) %>% 
  group_by(city, pollutant, six_days_period) %>% 
  mutate(
    id_period = (as.integer(date) + rand) %% 6,
    missing_crea = !(id_period == 1)
  ) %>% 
  ungroup() %>% 
  select(-six_days_period, -id_period) 
```

# Correlation with covariates

In order to create the most realistic pattern of missingness, we want to estimate the probability of missingness as a function of the other covariates.

```{r}
mod <- glm(data = data_daily, 
          missing ~ rainfall_height + rainfall_duration, family = binomial(link = "probit"))
```


```{r}
pred <- predict(mod, type="response")

t <- data_daily %>% 
  filter(!is.na(rainfall_height) & !is.na(rainfall_duration) & !is.na(missing)) %>% 
  mutate(missing_crea = rbernoulli(length(pred), p = pred))
```











