---
title: "Power Simulation Exercise: Standard Model"
description: |
  Marseille Data 2008-2018.
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
  - name: Léo Zabrocki 
    url: https://www.parisschoolofeconomics.eu/en/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

In this document, we carry out a simulation exercise measure the statistical power of a standard analysis of the short-term effects of air pollution on health. We use here data from Marseille over the 2008-2018 period.

# Loading and Formatting Data

We load the packages:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualisation
library(modelr) # modelling within the tidyverse
library(retrodesign) # formulas for type-m and type-s errors
library(knitr) # for tables
```

We load our custom ggplot2 theme:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
source(here::here("2.scripts", "script_custom_ggplot_theme.R"))
```

We open the data:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- readRDS(here::here("1.data", "1.standard_model", "data_marseille_daily_2008_2018.rds"))
```

We transform some calendar variables as factors:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate_at(vars(year, holidays_dummy, bank_day_dummy), ~ as.factor(.))
```

We select the relevant period as emergency admissions data are available from 2010:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  filter(!(year %in% c("2008", "2009")))
```

We create wind direction categories:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate(wind_direction_categories = cut(wind_direction, breaks = seq(0, 360, by  = 90), include.lowest = TRUE) %>%
  recode(., "[0,90]" = "North-East", "(90,180]" = "South-East", "(180,270]" = "South-West", "(270,360]" = "North-West"))
```

We create the one-day lag for wind speed and wind direction:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate(wind_speed_lag_1 = lag(wind_speed),
         wind_direction_categories_lag_1 = lag(wind_direction_categories))
```

We create a rainfall dummy:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate(rainfall_dummy = ifelse(rainfall_height>0, "True", "False"))
```

We also need to compute the the average mean over the current and previous days for continuous dependent variables

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate_at(vars(mean_no2_agregate, mean_o3_l, mean_pm10_agregate, mean_pm25_l, temperature_average, humidity_average), list("01" = ~ zoo::rollmean(., k = 2, align = "right", fill = NA)))
```

We then scale the new pollutant measures:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate_at(vars(mean_no2_agregate_01:humidity_average_01), ~ scale(.))
```

We take the logarithm of the `emergency_cv_r` variable to measure a percentage change in this outcome:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  mutate(emergency_cv_r = log(emergency_cv_r))
```

We finally select relevant variables:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data <- data %>%
  select(emergency_cv_r, mean_no2_agregate, temperature_average_01, humidity_average_01, rainfall_dummy, wind_speed, wind_speed_lag_1, wind_direction_categories, wind_direction_categories_lag_1, weekday, holidays_dummy, bank_day_dummy, month, year)
```      

# Simulations

### Preparing the simulations

We first want to have an idea of the effect of `mean_no2_agregate` on `emergency_cv_r` in our data:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
fitted_model_real_data <- lm(emergency_cv_r ~ mean_no2_agregate + 
              temperature_average_01 + I(temperature_average_01^2) +
              humidity_average_01 +
              rainfall_dummy +
              wind_speed + wind_speed_lag_1 + 
              wind_direction_categories + wind_direction_categories_lag_1 +
              weekday + holidays_dummy + bank_day_dummy + month*year, data = data)

fitted_model_real_data %>%
  broom::tidy(., conf.int = TRUE) %>%
  filter(term == "mean_no2_agregate")
```      

We see that a one standard deviation in `mean_no2_agregate` lead to 0.05% (95% CI: [-0.035; 0.139]) increase in `emergency_cv_r`.

Using our data and this model, we can simulate fake-data for different effect sizes of `mean_no2_agregate` on
`emergency_cv_r`. We first create a nested tibble where store our data, variables recording the effect sizes (effect sizes = 0.05%, 0.1% , 0.5% and 1%) and the sample size (N = 1068, 1780, 2492, 3650) and the fit of the true model.

```{r, echo=TRUE, message = FALSE, warning = FALSE}
data_simulations <- data %>% 
  nest(data = everything()) %>%
  crossing(effect_size = c(0.0005, 0.001, 0.005, 0.01)) %>%
  crossing(sample_size = c(1068, 1780, 2492, 3650)) %>%
  mutate(true_model = map(data, ~ lm(emergency_cv_r ~ mean_no2_agregate + 
                                       temperature_average_01 +
                                       I(temperature_average_01^2) +
                                       humidity_average_01 +
                                       rainfall_dummy +
                                       wind_speed + wind_speed_lag_1 + 
                                       wind_direction_categories + wind_direction_categories_lag_1 +
                                       weekday + holidays_dummy + bank_day_dummy + month*year, data = .)))
```      

Once we have fitted the true model on our data, we need to modify the effect size of `mean_no2_agregate` in order to create new fake `emergency_cv_r` observations:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
function_model_effect_size <- function(model, effect_size){
  model$coefficients[2] <- effect_size
  return(model)
} 

data_simulations <- data_simulations %>%
  mutate(true_model = map2(true_model, effect_size, ~ function_model_effect_size(.x, .y)))
```        

We finally need to retrieve the mean and the standard deviation of residuals found for the `fitted_model_real_data` to add some noise in our simulations:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
# get the distribution of residuals
mean_residuals <- mean(residuals(fitted_model_real_data))
sd_residuals <- sd(residuals(fitted_model_real_data))
```      

### Function to create fake-data and run the model

We create a function to create fake-data, run on them the model for which know the true effect size and retrieve the estimate and p-value for `mean_no2_agregate`. The function takes three arguments:the data, a sample size, and a model:

```{r, eval = FALSE, message = FALSE, warning = FALSE}
function_power_simulation <- function(data, sample_size, true_model){
simulation_results <- data %>%
  sample_n(., sample_size, replace = TRUE) %>%
  add_predictions(true_model, var = "predicted_emergency_cv_r") %>%
  mutate(predicted_emergency_cv_r = predicted_emergency_cv_r + rnorm(nrow(.), mean_residuals, sd_residuals)) %>%
  lm(predicted_emergency_cv_r ~ mean_no2_agregate + 
       temperature_average_01 + I(temperature_average_01^2) +
       humidity_average_01 +
       rainfall_dummy +
       wind_speed + wind_speed_lag_1 + 
       wind_direction_categories + wind_direction_categories_lag_1 +
       weekday + holidays_dummy + bank_day_dummy + month*year, data = .) %>%
  broom::tidy() %>%
  filter(term == "mean_no2_agregate") %>%
  select(estimate, p.value)
  
  return(simulation_results)
}
```      

### Running and Cleaning the Simulations

We run the simulations:

```{r, eval = FALSE, message = FALSE, warning = FALSE}
data_simulations <- data_simulations %>%
  crossing(id_sim = seq(1:1000)) %>%
  mutate(simulation_results = pmap(list(data, sample_size, true_model), function_power_simulation))
``` 

We compute the power, type m and s errors:

```{r, eval = FALSE, message = FALSE, warning = FALSE}
data_simulation_results <- data_simulations %>%
  select(-data, - true_model) %>%
  unnest(simulation_results)

data_simulation_results_power <- data_simulation_results %>%
  group_by(sample_size, effect_size) %>%
  summarise(power = mean(p.value<= 0.05)*100)

data_simulation_results_m_error <- data_simulation_results %>%
  group_by(sample_size, effect_size) %>%
  filter(p.value <= 0.05) %>%
  summarise(type_m_error = mean(abs(estimate)/effect_size))

data_simulation_results_s_error <- data_simulation_results %>%
  group_by(sample_size, effect_size) %>%
  filter(p.value <= 0.05) %>%
  summarise(type_s_error = (sum(estimate<0)/n())*100)

# join the figures together
data_simulation_results_all <- left_join(data_simulation_results_power, data_simulation_results_m_error, by = c("sample_size", "effect_size")) %>%
  left_join(., data_simulation_results_s_error, by = c("sample_size", "effect_size"))
``` 


### Simulation Results

We plot below the simulations results:

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width=20, fig.height=8, fig.fullwidth=TRUE, dev = "CairoPNG"}
# load simulation results data
data_simulation_results <- readRDS(here::here("1.data", "1.standard_model", "simulation_results.RDS"))

# make the graph
graph <- data_simulation_results %>%
  mutate_at(vars(sample_size, effect_size), ~ as.factor(.)) %>%
  mutate(effect_size = case_when(effect_size == "5e-04" ~ "+0.05%",
                                 effect_size == "0.001" ~ "+0.1%",
                                 effect_size == "0.005" ~ "+0.5%",
                                 effect_size == "0.01" ~ "+1%")) %>%
  rename("Power (%)" = power, "Type M Error (Bias)" = type_m_error, "Type S Error (%)" = type_s_error) %>%
  pivot_longer(cols= c(`Power (%)`:`Type S Error (%)`), names_to = "variable", values_to = "value") %>%
  ggplot(., aes(x = sample_size, y = value, group = effect_size, color = effect_size)) +
  geom_line(linetype = "dashed") +
  geom_point(size = 4) +
  scale_color_manual(values=c("#f2cc8f", "#3d405b", "#81b29a", "#e07a5f")) +
  facet_wrap(~ variable, scales = "free", nrow = 1) +
  xlab("Sample Size") + ylab("Value") +
  ggtitle("Power Simulations Results", subtitle = "The effect of a one standard deviation increase in NO2 on Emergency Admissions.") +
  labs(color = "Effect Size:") +
  custom_theme +
  theme(legend.position = "top", legend.justification = "left", legend.direction = "horizontal")

# print graph
graph

# save graph
ggsave(graph, filename = here::here("3.outputs", "1.figures", "graph.pdf"), 
       width = 45, height = 20, units = "cm", device = cairo_pdf)
``` 

We can compare our results with those of `retrodesign` package:

```{r, eval = TRUE, message = FALSE, warning = FALSE}
tibble(effect_size = c(0.0005, 0.001, 0.005, 0.01), standard_error = c(rep(0.00045, 4))) %>%
  mutate(power = map2(effect_size, standard_error, ~ retro_design(.x, .y)$power*100),
         type_s = map2(effect_size, standard_error, ~ retro_design(.x, .y)$typeS*100),
         type_m = map2(effect_size, standard_error, ~ retro_design(.x, .y)$typeM)) %>%
  unnest(cols = c(power, type_s, type_m)) %>%
  mutate_at(vars(power:type_m), ~ round(., 1)) %>%
  mutate(effect_size = case_when(effect_size == "5e-04" ~ "+0.05%",
                                 effect_size == "0.001" ~ "+0.1%",
                                 effect_size == "0.005" ~ "+0.5%",
                                 effect_size == "0.01" ~ "+1%")) %>%
  kable()
``` 



