---
title: "Power Simulation Exercise: Standard Model"
description: |
  Marseille Data 2008-2018.
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
  - name: Léo Zabrocki 
    url: https://www.parisschoolofeconomics.eu/en/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

In this document, we carry out a simulation exercise measure the statistical power of a standard analysis of the short-term effects of air pollution on health. We use here data from Marseille over the 2008-2018 period.

# Loading and Formatting Data

We load the packages:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(modelr) # modeling within the tidyverse
library(retrodesign) # formulas for type-m and type-s errors
library(knitr) # for tables
library(mediocrethemes)

set_mediocre_all()
```

# Running a Very Simple Power Simulation

### Set-Up of the Simulations

We create a science table where we observe the two potential outcomes - here the mortality counts under a control and hypothetical treatment equal to +0.42 death - for a city over 10 000 days:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
# create the science table
data <- 
  tibble(
    y_0 = rpois(10000, 42), 
    y_1 = y_0 + rnorm(10000, mean = 0.42, sd = 0.5),
    treatment = c(rep(0, 5000), rep(1, 5000))
  )
# display the science table
head(data)
```

We added to this science table a treatment indicator that we are going to randomly permute to carry out many randomized experiments for various sample sizes. We define below a function to randomly sample a given number of days, randomize the treatment and compute the average difference in mortality counts between treated and control units:

```{r, echo=TRUE, message = FALSE, warning = FALSE}
# power simulation function
function_power_simulation <- function(data, sample_size){
simulation_results <- data %>%
  sample_n(., sample_size, replace = TRUE) %>%
  mutate(
    permuted_treatment = sample(treatment, replace = FALSE),
    mortality_count = y_0 * (1 - permuted_treatment) + y_1 * permuted_treatment
  ) %>%
  lm(mortality_count ~ permuted_treatment, data = .) %>%
  broom::tidy() %>%
  filter(term == "permuted_treatment") %>%
  select(estimate, p.value)

  return(simulation_results)
}
```

### Running the Simulations

We run the simulations:

```{r, eval = FALSE, message = FALSE, warning = FALSE}
data_simulations <- data %>%
  nest(data = everything()) %>%
  crossing(sample_size = c(100, 250, 500, 750, 1000, 1500, 2500, 5000)) %>%
  mutate(
    simulation_results = map2(
      data, 
      sample_size, 
      ~ rerun(2500, function_power_simulation(.x, .y))
    )
  )

saveRDS(data_simulations, "../Outputs/data_simulations.RDS")
``` 

We compute the power, type m and s errors:

```{r message = FALSE, warning = FALSE}
data_simulations <- readRDS("../Outputs/data_simulations.RDS")

# we transform lists of results into data frames with 100 rows 
# and two columns for the coefficients and p-values:
data_simulation_results <- data_simulations %>%
  select(-data) %>%
  unnest(simulation_results) %>% 
  unnest(simulation_results)

data_simulation_results_all <- data_simulation_results %>%
  group_by(sample_size) %>%
  mutate(power = mean(p.value<= 0.05)*100) %>% 
  filter(p.value <= 0.05) %>%
  summarise(
    power = unique(power),
    type_m_error = mean(abs(estimate)/0.42), 
    type_s_error = (sum(estimate<0)/n())*100
  )
``` 


### Simulation Results

We plot below the simulations results:

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width=20, fig.height=8, fig.fullwidth=TRUE, dev = "CairoPNG"}
# make the graph
graph <- data_simulation_results_all %>%
  mutate_at(vars(sample_size), ~ as.factor(.)) %>%
  rename("Power (%)" = power, "Type M Error (Bias)" = type_m_error, "Type S Error (%)" = type_s_error) %>%
  pivot_longer(cols= c(`Power (%)`:`Type S Error (%)`), names_to = "variable", values_to = "value") %>%
  ggplot(., aes(x = sample_size, y = value, group = "l")) +
  geom_line(linetype = "dashed") +
  geom_point() + 
  # geom_point(shape = 21, size = 4, fill = "coral", color = "black") +
  facet_wrap(~ variable, scales = "free", nrow = 1) +
  xlab("Sample Size") + ylab("") +
  ggtitle("Power Simulations Results") 

# print graph
graph

# save graph
# ggsave(graph, filename = here::here("3.outputs", "1.figures", "graph_demo_power.pdf"), 
#        width = 45, height = 20, units = "cm", device = cairo_pdf)
``` 



