---
title: "Systematic literature review"
subtitle: "Missing data in studies of health effects of air pollution"
output: 
  html_notebook: 
    toc: yes
    theme: simplex
    highlight: pygments
  html_document:
    theme: simplex
    highlight: pygments
author: "Vincent Bagilet, Léo Zabrocki"
date: "October 30, 2020"
---

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE,
               message = FALSE,
               warning = FALSE)  
```  

```{r include=FALSE}
library(tidyverse)
library(fulltext)
```

## Purpose of the document

In the present document, we aim to conduct a systematic review of the literature on health effects of air pollution. We carry out an extremely simple text mining analysis of this literature in order to have a sense of the proportion of articles studying health effects of air pollution data taking the issue of missing data into account.

## Retreiving article texts

To carry out this analysis, we use the [`fulltext`](https://books.ropensci.org/fulltext/) package to get the full text of each article whose title contains the words "air pollution" and "health". This set of articles might be too restrictive and we may want to broaden our scope later. 

To begin with, we focus on articles published on Scopus (one needs to register to get an API key).


```{r}
#set folder to store downloads
cache_options_set(full_path = "/Users/vincentbagilet/Documents/fulltext")

#ft_search() to run a search
search <- ft_search("air pollution health", from = "scopus", limit = 3)

#retreive the full text
search$plos$data$id %>% 
  ft_get(progress = TRUE) 

#put text into a table format
with_text <- ft_table() #searches for documents in the cache directory 

with_text %>% 
  mutate(
    contains_missing = str_detect(text, pattern = "missing")
  )

#I need to pipe all this of course but I am just recording this so that I do not forget
```

Once we have accessed files





