---
title: "Systematic literature review"
author:
- name: Vincent Bagilet
  url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
  affiliation: Columbia University
  affiliation_url: https://www.columbia.edu/
- name: Léo Zabrocki
  url: https://www.parisschoolofeconomics.eu/en/
  affiliation: Paris School of Economics
  affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
output:
  distill::distill_article: default
  html_notebook: default
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = FALSE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

```{r include=FALSE}
library(tidyverse)
library(fulltext)
library(tidytext)
library(wordcloud)
library(retrodesign)
library(mediocrethemes)
library(lubridate)
library(DT)

set_mediocre_all()
set.seed(1)
```

# Purpose of the document

In the present document, we aim to conduct a systematic review of the literature on short term health effects of air pollution. The objective is two fold: 
- Retrieve effect sizes and confidence intervals in order to compute power, type M and type S error in the literature
- Get a sense of the proportion of papers in this literature discussing power and missing data issues.  

# Motivation

In this section, we discuss the importance of such an analysis.

# Power analysis

In this section, we implement robustness tests in order to compute the power, type M and type S error in the studied articles. We look at what would be the power, type M and type S error if the true effect was a fraction of the measured effect. We retrieved estimates and confidence intervals of articles in the literature of interest in [another document](systematic_lit_review_geting_abstracts.html). Before looking into the power analysis itself, we look at the characteristics of the articles considered.

## Articles characteristics

### Full set of articles

We retrieved the articles using the following query:

'TITLE(("air pollution"  OR "air quality" OR "particulate matter" OR ozone OR "nitrogen dioxide" OR "sulfur dioxide") AND ("emergency" OR "mortality") AND NOT ("long term" or "long-term")) AND ("particulate matter" OR ozone OR "nitrogen dioxide" OR "sulfur dioxide")'

```{r}
articles_effect <- readRDS("../Outputs/articles_effect.RDS")
metadata_lit_review <- readRDS("../Outputs/metadata_lit_review.RDS")
abstracts <- readRDS("../Outputs/abstracts.RDS")
```

This query returns `r metadata_lit_review %>% distinct(title) %>% nrow()`. Based on the abstracts, we can briefly explore the main (unsurprising) themes of the articles.

```{r}
abstracts %>%
  unnest_tokens(word, abstract, to_lower = TRUE) %>% 
  anti_join(tidytext::stop_words, by = "word") %>%
  count(word) %>%
  with(wordcloud::wordcloud(word, n, max.words = 100, random.color = TRUE, colors = "#00313C"))
```

```{r fig.asp = 1}
abstracts %>%
  unnest_tokens(word, abstract, to_lower = TRUE) %>% 
  anti_join(tidytext::stop_words, by = "word") %>%
  count(word, sort = TRUE) %>%
  filter(n > 1000) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL, title = "Main themes in the abstracts", subtitle = "Droping usual stop words") 
```

### Abstracts with effects and confidence intervals

```{r}
string_confint <- "((?<!(\\.|\\d))95\\s?%|\\b95\\s(per(\\s?)cent)|\\bCI(s?)\\b|\\bPI(s?)\\b|\\b(i|I)nterval|\\b(c|C)onfidence\\s(i|I)nterval|\\b(c|C)redible\\s(i|I)nterval|\\b(p|P)osterior\\s(i|I)nterval)"

abstracts_CI <- articles_effect %>% 
  mutate(
    contains_CI = str_detect(abstract, string_confint)
  ) %>%
  group_by(title) %>% 
  mutate(has_effect = (sum(!is.na(effect), na.rm = TRUE) > 0)) %>% 
  ungroup() %>% 
  select(-effect, -low_CI, -up_CI) %>% 
  distinct()
```

Out of all articles returned by the query, `r articles_contains_CI %>% distinct(title) %>% nrow()` display confidence intervals. "CI", "confidence interval", etc. These articles are the following:

```{r}
abstracts_CI %>% 
  filter(contains_CI) %>% 
  select(authors, pub_date, title, journal) %>% 
  datatable()
```

In these articles, we retrieve valid effects and confidence intervals in the following proportions:^[Note that a bunch of abstracts contain the phrase "CI" without actually displaying effects and confidence intervals.]

```{r}
abstracts_CI %>%
  filter(contains_CI) %>% 
  count(has_effect) %>% 
  mutate(
    prop = n/sum(n),
    has_effect = ifelse(has_effect, "Yes", "No")
  ) %>% 
  arrange(desc(has_effect)) %>% 
  kable(col.names = c("Effect retreived", "Number of articles", "Proportion"), caption = "Number of articles for which at least one effect is retrieved (out of those containing the phrase CI)")
```

This corresponds to `r articles_contains_CI %>% filter(!is.na(effect)) %>% nrow()` valid effects and associated confidence intervals.

### Comparison between abstracts with and without detected effects

In this subsection, we investigate whether there are systematic differences between articles displaying an effect in the abstract (effect that we detected) and articles that do not. 

We first wonder whether there are disparities in publication dates. It might be the case that displaying effects in the abstract was a feature of a given period.

```{r}
abstracts_CI %>% 
  mutate(has_effect = ifelse(has_effect, "Effect detected", "Effect not detected")) %>% 
  ggplot() +
  geom_density(aes(x = year(pub_date), fill = has_effect, , color = has_effect)) +
  labs(title = "Distribution of publication year", subtitle = "Comparison between articles for which an effect is retrieved and not", x = "Publication year", y = "Density") + 
  scale_fill_discrete(name = "")
```
Even though there are slightly more recent (2010-2020) articles for which effects are retrieved, the difference does not seem to be substantial.

We also investigate whether there are differences in the journals in which the articles are published.

```{r fig.asp=1}
abstracts_CI %>% 
  mutate(has_effect = ifelse(has_effect, "Effect detected", "Effect not detected")) %>% 
  group_by(has_effect) %>% 
  mutate(nb_has_effect = n()) %>% 
  count(journal, nb_has_effect, has_effect, sort = TRUE) %>% 
  mutate(prop = n/nb_has_effect) %>% 
  select(-nb_has_effect, -n) %>% 
  group_by(journal) %>% 
  mutate(keep = (sum(prop) > 0.02)) %>% 
  filter(keep) %>% 
  select(-keep) %>% 
  ungroup() %>% 
  mutate(journal = reorder(journal, prop)) %>% 
  ggplot() +
  geom_col(aes(prop, journal, fill = has_effect)) +
  # facet_wrap(~ has_effect) +
  labs(y = NULL, title = "Main themes in the abstracts", subtitle = "Droping usual stop words") +
  scale_fill_discrete(name = "")
```

For this analysis to be informative, we would need to cluster the journals into groups (*eg* epidemiology journals, general science journals, etc).

Then, we wonder if the the themes considered in each types of abstracts differ.

```{r fig.asp=1}
abstracts_CI %>%
  unnest_tokens(word, abstract, to_lower = TRUE) %>% 
  anti_join(tidytext::stop_words, by = "word") %>%
  count(word, has_effect, sort = TRUE) %>%
  group_by(word) %>% 
  mutate(keep = (sum(n) > 1200)) %>% 
  filter(keep) %>% 
  select(-keep) %>% 
  mutate(has_effect = ifelse(has_effect, "Effect detected", "Effect not detected")) %>% 
  group_by(has_effect) %>% 
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word, fill = has_effect)) +
  geom_col(position = "dodge") +
  # facet_wrap(~ has_effect) + 
  labs(y = NULL, title = "Main themes in the abstracts", subtitle = "Comparison between articles with and without detected effect") +
  scale_fill_discrete(name = "")
```

Apart from a few key terms, such as CI, 95 for instance, there are no huge variations in the themes. 

Now that we have quickly compared the articles for which we retrieve an effect an those for which we don't, we can dig further into the analysis of the estimates retrieved.

## Power analysis

To carry out our power analysis, we use the package `retrodesign` which computes post analysis design calculations (power, type M and type S errors). To use the `retro_design()` function, we first need to compute the standard error of the estimate. Probably due to rounding effect, we often do not get the same value for the standard error if we compute it using the upper or the lower bound of the CI. Thus, we average across the two values obtained.

```{r}
articles_retro <- articles_effect %>% 
  mutate(
    se_up = (up_CI - effect)/1.96,
    se_low = (low_CI - effect)/(-1.96),
    se = (se_up + se_low)/2
  ) %>% 
  select(-se_low, -se_up) %>% 
  filter(!is.na(effect)) %>% 
  mutate(
    # retro_0.01 = as_tibble(retro_design(effect*0.01, se)),
    # retro_0.05 = as_tibble(retro_design(effect*0.05, se)),
    # retro_0.1 = as_tibble(retro_design(effect*0.1, se)),
    retro_0.1 = as_tibble(retro_design(effect*0.33, se)),
    retro_0.5 = as_tibble(retro_design(effect*0.5, se)),
    # retro_0.67 = as_tibble(retro_design(effect*0.67, se)),
    retro_0.75 = as_tibble(retro_design(effect*0.75, se)),
    # retro_0.9 = as_tibble(retro_design(effect*0.9, se)) 
  ) %>% 
  pivot_longer(
    starts_with("retro"), 
    names_to = "prop_true_effect", 
    values_to = "computed"
  ) %>% 
  mutate(
    power = computed$power,
    typeS = computed$typeS,
    typeM = computed$typeM,
    prop_true_effect = as.numeric(str_sub(prop_true_effect, 7, nchar(prop_true_effect)))
  ) %>% 
  select(-computed)
```

## Results and graphs

Then, we quickly explore the results. First, we compute the average and median power, type M and type S errors.

```{r}
articles_retro %>% 
  filter(typeM < Inf) %>% 
  group_by(prop_true_effect) %>% 
  summarise(
    mean_power = mean(power, na.rm = TRUE),
    mean_typeM = mean(typeM, na.rm = TRUE),
    mean_typeS = mean(typeS, na.rm = TRUE),
    median_power = median(power, na.rm = TRUE),
    median_typeM = median(typeM, na.rm = TRUE),
    median_typeS = median(typeS, na.rm = TRUE)
  ) %>% 
  kable()
```


Then, we look at the distribution of power, type M and type S error across simulation and for different size of true effect.

```{r}
articles_retro %>% 
  ggplot(aes(x = power)) + 
  geom_histogram(bins = 10) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Distribution of power in retrodesign simulations", subtitle = "If the size of true effect was a fraction of the measured effect")
  # geom_density()

articles_retro %>% 
  ggplot(aes(x = typeM)) + 
  geom_histogram(bins = 10) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Distribution of type M error in retrodesign simulations", subtitle = "If the size of true effect was a fraction of the measured effect") +
  scale_x_continuous(trans='log10')

articles_retro %>% 
  ggplot(aes(x = typeS)) + 
  geom_histogram(bins = 10) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Distribution of type S error in retrodesign simulations", subtitle = "If the size of true effect was a fraction of the measured effect")
```

Then, we look at the relation between power, type M and type S error and true effect size.

```{r}
articles_retro %>% 
  ggplot() +
  geom_point(aes(x = power, y = typeM)) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Link between type M and power in retrodesign simulations", subtitle = "If the size of true effect was a fraction of the measured effect", x = "Power", y = "Type M")

articles_retro %>% 
  ggplot() +
  geom_point(aes(x = power, y = typeS)) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Link between type S and power in retrodesign simulations", subtitle = "If the size of true effect was a fraction of the measured effect", x = "Power", y = "Type S")

articles_retro %>% 
  filter(typeM < Inf) %>% 
  group_by(prop_true_effect) %>% 
  summarise(typeM = mean(typeM, na.rm = TRUE)) %>% 
  ggplot() +
  geom_point(aes(x = prop_true_effect, y = typeM)) +
  labs(title = "Link between type M and 'true effect' in retrodesign simulations", x = "True effect as a proportion of the measured effect", y = "Type M")

articles_retro %>% 
  group_by(prop_true_effect) %>% 
  summarise(typeS = mean(typeS, na.rm = TRUE)) %>% 
  ggplot() +
  geom_point(aes(x = prop_true_effect, y = typeS)) +
  labs(title = "Link between type S and 'true effect' in retrodesign simulations", x = "True effect as a proportion of the measured effect", y = "Type S")
```

We can also look at how power, type M and type S error evolved with publication date.

```{r}
articles_retro %>% 
  group_by(year = year(pub_date), prop_true_effect) %>%
  mutate(power = mean(power, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = year, y = power)) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Evolution of power with publication date", x = "Publication date", y = "Power")

articles_retro %>% 
  group_by(year = year(pub_date), prop_true_effect) %>%
  mutate(typeM = mean(typeM, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = year, y = typeM)) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Evolution of type M with publication date", x = "Publication date", y = "Type M")

articles_retro %>% 
  group_by(year = year(pub_date), prop_true_effect) %>%
  mutate(typeS = mean(typeS, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = year, y = typeS)) +
  facet_wrap(~ prop_true_effect) +
  labs(title = "Evolution of type S with publication date", x = "Publication date", y = "Type S")
```

# Analysis of papers dealing with power and missingness issues

## Retreiving the full texts

We then download the full texts (using the `ft_get` functions). The full texts are stored in an xml format in the cache directory. Note that due to my IP address being located outside of Columbia, I cannot access the texts from Scopus.

```{r eval = FALSE}
metadata_scopus$doi %>% 
  ft_get(progress = TRUE)

metadata_entrez$doi %>% 
  ft_get(from = "entrez", progress = TRUE, type = "pdf")
```

Once we have downloaded all the files, we can put them into a table format, before analyzing them.

```{r eval = FALSE}
#put text into a table format
article_texts <- ft_table() %>% #searches for documents in the cache directory 
  select(-paths, -ids_norm) %>% 
  .$text 

saveRDS(article_texts, "../Outputs/article_texts.RDS")
```

## Analysis

To analyse the texts, we first start by simply exploring the proportion of articles mentioning the words "missing" and "power".

```{r eval = FALSE}
readRDS("../Outputs/article_texts.RDS")

article_texts %>%
  summarise(
    contains_missing = sum(str_detect(text, pattern = "\\bmissing"))/n(), #could put a vector with various terms
    contains_power = sum(str_detect(text, pattern = "\\bpower\\b"))/n() #statistical power
  )
```

<!-- # Test with rentrez package -->

<!-- ```{r} -->
<!-- querry_entrez <- "((air pollution[TITL]  OR air quality[TITL] OR particulate matter[TITL] OR ozone[TITL] OR nitrogen dioxide[TITL] OR sulfur dioxide[TITL]) AND (emergency[TITL] OR mortality[TITL])) AND (particulate matter[TIAB] OR ozone[TIAB] OR nitrogen dioxide[TIAB] OR sulfur dioxide[TIAB])" -->

<!-- rentrez_search <- entrez_search(db = "pubmed", term = querry_entrez, use_history = TRUE) -->
<!-- rentrez_ids <- rentrez_search$ids %>% as.numeric() -->

<!-- rentrez_texts <- entrez_fetch(db = "pubmed", id = rentrez_ids, rettype = "xml") -->

<!-- t <- rentrez_texts %>% -->
<!--   xmlParse() %>% -->
<!--   xmlToList() -->

<!-- retrieve_abstract <- function(lis) { -->
<!--   df <- lis$MedlineCitation$Article$Abstract %>% -->
<!--   map_df(., data.frame) %>% -->
<!--   as.tibble() %>% -->
<!--   distinct() -->

<!--   # %>% -->
<!--   # filter(!is.na(text)) %>% -->
<!--   # select(-3) -->

<!--   return(df) -->
<!-- } -->

<!-- out <- map_df(t, retrieve_abstract) -->

<!-- r <- t$PubmedArticle$MedlineCitation$Article$Abstract %>% -->
<!--   map_df(., data.frame) %>% -->
<!--   as.tibble() %>% -->
<!--   distinct() %>% -->
<!--   filter(!is.na(text)) %>% -->
<!--   select(-3) -->

<!-- t$PubmedArticle$MedlineCitation$Article -->
<!-- ``` -->




